Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-08 writing/embedded_meta_learning.tex
Starting section on meta-mappings

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks. For example, once a system has been trained to play a game, it should be able to alter its strategy in accordance with a change in task demands, such as trying to lose the game, or trying to play without using a certain action. The challenge is that the system must adapt to performing these tasks without seeing any data from the new task, only from either examples of similar transformations on different tasks (``here's how to change your go strategy to a losing one, here's how to change your checkers strategy to a losing one, now try to lose at chess'') or from natural language instructions (``try to lose at chess''). These mappings provide a new type of zero-shot transfer from one task to another -- the system must transform its behavior to adapt to a new task without ever having seen training data from that task. This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. \par
To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating task behaviors and data as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. By embedding task behaviors, data, and language in the same space, we can conceptually think of both basic task solutions and adaptive functions, such as following instructions or altering behavior, as mappings in function space. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute meta-mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits, such as a useful framework for continual learning. \par
\section{Embedded Meta-Learning (EML) architecture}
We propose a new class of Embedded Meta-Learning (EML) architectures which consist of two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior based on examples). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a braod enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par 
\begin{tikzpicture}[auto, scale=0.9]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

>>>>

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks. For example, once a system has been trained to play a game, it should be able to alter its strategy in accordance with a change in task demands, such as trying to lose the game, or trying to play without using a certain action. The challenge is that the system must adapt to performing these tasks without seeing any data from the new task, only from either examples of similar transformations on different tasks (``here's how to change your go strategy to a losing one, here's how to change your checkers strategy to a losing one, now try to lose at chess'') or from natural language instructions (``try to lose at chess''). These mappings provide a new type of zero-shot transfer from one task to another -- the system must transform its behavior to adapt to a new task without ever having seen training data from that task. This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. \par
To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating task behaviors and data as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. By embedding task behaviors, data, and language in the same space, we can conceptually think of both basic task solutions and adaptive functions, such as following instructions or altering behavior, as mappings in function space. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute meta-mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits, such as a useful framework for continual learning. \par

\section{Meta-mappings}
We propose the challenge of meta-mappings, that is, mappings over tasks. These can take several flavors, including mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been considered in prior work, the latter is more novel. \par
We argue that these meta-mappings are a useful way to think about the flexibility that humans have. For example, suppose an agent knows how to navigate from the airport to downtown. It should be able to adapt to changes of the environment such as street closures. This can be viewed as a change in the task being performed, from navigating through one map to navigating through another. Similarly, a system that is trained to play one level of a video game should be able to adapt to another level where the enemies change color. These again can be seen as two related tasks. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to use. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not capture the extent of human flexibility. In particular, we can use our knowledge of a task to adapt to these type of task alterations zero-shot, that is, without seeing a single example from the new task. \par

\section{Embedded Meta-Learning (EML) architecture}
We propose a new class of Embedded Meta-Learning (EML) architectures which consist of two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior based on examples). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a braod enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.9]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};
